
\section{ProofOptimizer: LLMs for Proof Simplification}

\subsection{Training}

% \smallsec{Lean Base Model} First, we train a general-purpose Lean model by fine-tuning \texttt{Qwen-2.5-7B-Instruct} on a combination of five tasks: natural language problem solving, Lean 4 code completion, auto-formalization (problems and solutions), formal theorem proving, and tactic/proof state prediction. 

% \smallsec{Dataset for Proof Simplification} We employ a four-stage pipeline to generate high-quality proof simplification training data.
% \begin{enumerate}[leftmargin=0.7cm, topsep=0pt]
% \itemsep1pt 
%     \item \textit{Problem Collection}: We first compile a dataset of theorem proving problems from \texttt{Goedel-Pset}, filtering out simple computational problems. Each problem consists of a natural language problem, solution, and Lean problem statement.
%     \item \textit{Proof Sketching}: We train a model that formalizes a problem's natural language solution into a Lean proof sketch consisting of a few high-level proof steps (usually 2-10) with lower level details omitted and filled in with Lean's \texttt{sorry} tactic.
%     \item \textit{Theorem Extraction and Filtering}: For each proof sketch, we extract each proof step into its own separate theorem. At the core, we are taking longer proofs and breaking them down into separate sub-theorems. We collect a total of 518K theorems this way. As we found some of these theorems to be trivial, we design an automation tactic to filter these out, leaving 307K theorems remaining.
%     \item \textit{Proof Generation}: We use \texttt{Goedel-Prover-V2-32B} to generate proofs of these theorems. The model successfully produces Lean proofs of 145K theorems, which we use as our dataset for training.
% \end{enumerate}

% For more details about our base model and dataset collection, see Appendix \ref{appendix:base-model-and-data}. Next, we describe our two training recipes: expert iteration and online reinforcement learning.

\textbf{ProofOptimizer-ExpIt: Expert Iteration}: We leverage a STaR-like \citep{zelikman2022star} iterative training algorithm to improve our model. We start with our base model $\pi_0$ and a collection of 145K proofs $P_0$. At each iteration, we attempt to simplify each proof, train our model on successful proof simplifications, and use the collection of simplified proofs as seed proofs for the next iteration. 

More precisely, we have a 3-stage process. In the \textbf{sample} phase, for each proof $x \in P_i$, sample $4$ simplifications $Y_p \triangleq \{y_x^{1}, y_x^{2}, y_x^{3}, y_x^{4}\} \sim \pi_i(x)$. Then, in the \textbf{filter} phase, we use the Lean compiler to find the shortest correct simplification $y_x \in \{x\} \cup Y_x$. Create a training dataset of proof simplifications $D_i = \{(x, y_x) \mid \text{len}(y_x) \le 0.8 \cdot \text{len}(x), x \in P_i\}$. The length constraint is designed to encourage the model to learn more substantial simplifications rather than trivial ones. For iterations after the first, as $x$ may have been simplified from a more complex proof $x' \in P_0$, we also add $(x', y_x)$ pairs to $D_i$, which are valid and larger proof simplifications. Also, collect simplified proofs $\pi_{i+1} = \{s_x \mid x \in P_i\}$ for the next iteration. Finally, in the \textbf{train} stage, we fine-tune $\pi_i$ on $D_i$ to get $\pi_{i+1}$.

\textbf{ProofOptimizer-RL: Online Reinforcement Learning}: We also train a proof optimizer model with online RL. Using the same dataset as in expert iteration, the RL task consists of producing a valid but shorter proof $y$ for a statement given an initial proof $x$. The reward is the relative shortening $R(x, y) = \frac{|y| - |x|}{|x|}$
if $y$ is valid and $|y| \leq |x|$, and $R(x, y) = 0$ otherwise. We employ asynchronous GRPO \citep{shao2024grpo} with advantage $A_i = R_i - \frac{1}{k}\sum_{j \leq k} R_j$.

\subsection{Inference-Time Techniques} \label{subsec:inference-time-algorithms}
First, we implement a symbolic linter that removes extraneous tactics via Lean's \texttt{linter.unusedTactic} linter, which detects tactics that do not change the proof state and provides messages like \texttt{'norm\_num' tactic does nothing}. We compare two techniques. In \textbf{test-time RL}, we use the setup from training and perform RL on our two evaluation sets (jointly). Our test-time RL keeps the input proof fixed, meaning improvements occur solely in the model’s parameters. In \textbf{iterative proof shortening}, we first sample $k$ candidate shortenings and take the shortest correct one. Then, we sample $k$ shortenings of the new proof, take the shortest correct one -- and so on.