An evaluation is a lens through which we view the capabilities of a model. As
language models become increasingly capable at programming, mathematics, and
logical reasoning, the design of this lens becomes central to scientific
progress: a benchmark can reveal a capability that was previously hidden, but it
can also distort our understanding when it measures the wrong behavior or
contains confounding biases.

This thesis studies three lessons from evaluating language models on reasoning
tasks. First, focusing on the right metrics without bias is crucial. For code
and mathematical reasoning, final correctness alone can miss important distinctions:
a model may pass tests while producing brittle or low-quality code, solve an
inequality while giving an invalid proof, or appear strong because an evaluation
contains trivial negatives or contaminated examples. Second, looking beyond the score
reveals hidden nuances. Deeper analysis can surface structure that aggregate
performance conceals, such as differences between code generation and code
execution, gains from distillation that do not reflect deeper understanding, and
complementary failure modes between chain-of-thought prompting and neurosymbolic
reasoning.
Third, what we measure shapes what we build. Once evaluations become standard,
they define what counts as progress and influence the capabilities that future
systems are optimized to demonstrate.

Together, these lessons argue that evaluation should be treated not as a
secondary measurement step, but as a central part of building and understanding
AI systems for code and reasoning.
