\chapter*{Acknowledgments}
\addcontentsline{toc}{frontmatter}{Acknowledgments}

I feel incredibly lucky to have Armando as my advisor throughout my PhD. Armando always made sure I was enjoying myself, giving me the rare combination of complete freedom yet continual support. I never felt shy going to him about anything on my mind, whether it was about research papers, career decisions, or the most fun adventures to have at a conference destination. Armando's advice has always been honest yet encouraging. Early in my PhD, I asked Armando how to read papers. He didn't answer the question, but the advice he gave me was more useful than any direct answer he could have given: before reading any related work, I should sit with the problem myself for a while so I can have fresh ideas before being polluted. This advice shaped the way I did research moving forward: to develop my own taste and way of approaching problems instead of imitating what others had done. Since then, Armando has taught me much more: how to pick the right problems, how to develop ideas, and how to present my work. More than anything, he taught me how to trust myself and think differently. It's all because of Armando that these years has been as intellectually exciting as they have been enjoyable.

I am also very fortunate to have Jacob and Kaiyu on my committee. Through co-advising my MEng project, Jacob taught me a lot about both NLP and research more broadly. During meetings, he always asked me what the core question I wanted to answer was, beyond just have a cool demo or improving some benchmark numbers. This pushed me to always think deeper about fundamental research questions, something I always thought about in the years that followed.

Kaiyu introduced me to theorem proving, which later became the focus of my latter half of the PhD. Having worked and believed in the field long before it became mainstream, he showed me what it meant to continuously pursue a vision while embracing the changes of the time. Beyond research, Kaiyu has been a constant source of support, and he always gave me earnest advice, helping me navigate all kinds of situations during my PhD.

My brilliant mentors and friends at Meta FAIR helped me grow in all kinds of ways. Gabriel and Baptiste opened the doors to the AI CodeGen world. Sida encouraged me to always think creatively yet remain disciplined. Aram showed me what it means to do anything you can to unblock any obstacle in the way of your goals. I also grew a lot through all the other collaborators, mentors, and new friends I met at Meta, through research and life adventures alike: Amaury, Bartosz, Chris, Dmitrii, Edouard, Fabian, Francois, Gautier, Hugh, Kunhao, Luyu, Scott, Volker, Weijia, Yuxiang, Zacharias, and so many more.

Although I knew very little about ML systems at the time, the AMMO team at NVIDIA gave me the opportunity to grow and learn about the field. My mentor Huizi is one of the most ambitious people I know, and his ability to execute is just as impressive. Through him, I learned as much about ML systems as I learned about how to build and grow a startup.

I had a wonderful experience working at Amazon AWS with the Automated Reasoning Group. I'm lucky to have had Soonho and Leo as mentors, humble yet extremely talented. Through my time there, they showed me all the impactful ways formal methods were being used outside of an academic setting. I'm also grateful for all the other friends and colleagues I met there: Hongshan, Zhizhen, Brando, Andrew, Chenyang, Zhixing, Joseph, and Indraneel, all of whom are extremely passionate about AI for math and formal verification.

I would also like to thank all the mentors who took a chance on me in my naive undergraduate years. My very first research mentors at MIT, Costis and Dylan, introduced me to multi-agent RL, and I especially thank them for their patience when it was clear that they could have finished what I had done in under half the time. Lily and Luca guided me, pointed me to interesting directions in ML robustness, gave me more and more freedom, and constantly supported me with research advice.

In my third year of undergrad, I joined the OptiML group. I will never forget the first problem Suvrit fantasized that I would help him solve: proving a better bound of Polyak-Lojasiewicz functions. While I never achieved the final goal, he taught me how to persevere and progress through the hardest problems. I'm also glad to have had the opportunity to learn from the rest of the group: Alp, Chulhee, Hadi, Horia, Jingzhao, Josh, Kwangjun, Tiancheng, Xiang, and Yi. Everyone's passion for research and optimization was inspiring: group meetings were always a highlight of the week, and there was never an abundance of fun research problems to work on.

I thank Hima for introducing me to the world of explainability research, for giving me so much advice on my course project for almost a year after the class had ended, and for teaching me how to be a creative researcher.

I thank my MIT-IBM mentors Songtao and Parikshit for teaching me a lot about bilevel optimization, helping me grind through some of the nastiest proofs I've ever written in a paper, and repeatedly giving me hope when I felt like I was hitting a wall.

One of the best parts of graduate school is all the brilliant people I got to meet and learn from. First, I believe I am part of one of the happiest research groups at MIT. Everyone's interests are so diverse, and I've learned so many cool things I'd never even thought I wanted to learn: biology, robotics, cognitive science, programming languages, systems, and more. You all have brought much joy to my research experience: Ivan, Jimmy, Kavi, Kliment, Leo, Maddy, Megan, Michael, Nathan, Omar, Ramya, Ria, and Zed. Thanks to Jack for mentoring me during my MEng and teaching me a lot about databases and to Theo for being someone I could always talk to in the lab about AI for code ideas. Finally, I'm lucky to have had the opportunity to mentor so many talented undergraduate and masters students: Clarise, Daniela, Daniela, Ethan, Gilford, Haijia, Ivy, Jacob, Tamara, Tianyi, and Yeabsira.

Despite living hundreds of miles away when I started my PhD, Naman and Wen-Ding felt like second labmates to me. We always brainstormed the craziest ideas, updated each other on the latest, wrote many papers, and went on escapades during conferences and beyond. In early 2024, I learned that ACL 2024 was in Thailand, the perfect place to propose to my then-girlfriend. I was determined to make it happen. In under a month and a half, Naman, Wen-Ding, Theo and Celine helped me put together and submit a paper (did they know the real motivations?), and to this day I am eternally grateful. Throughout my PhD, the AI4Code community grew larger and larger, and I am fortunate to have many other friends to brainstorm and chat about code-related ideas (and struggles) from time to time: Dylan, Jenny, Kexun, Manish, Wenting, Zhaoyu, and Zora.

My many talented co-authors and collaborators gave me the chance to think about and work on on a variety of diverse projects: Alexis, Antoine, Ben, Cathy, Ced, Chanwoo, Diganta, Don, Gabe, Hung, Jiayi, Kaifeng, Pan, Pin-Yu, Rayan, Roger, Saiteja, Samy, Satyapriya, Shannon, Sirui, Terry, Tessa, Yijia, and Ziyang. You have all taught me so much, and I always look forward to being inspired by your ideas in our project meetings.

Finally, I thank my family, who was there long before I was, for constantly reminding me to chase my dreams. And finally, I thank my partner and love for always giving me the confidence and support that we can achieve anything together.
