\subsection{Model Details and Parameters} \label{sec:appendix_model_details}
We use a decoding temperature $T=0.8$ for all models. 
For \texttt{GPT-3.5} and \texttt{GPT-4}, we limit the maximum number of tokens to generate to 1024 for FOLIO and 4096 for ProofWriter (to accommodate for the previously mentioned larger number of premises involved in a typical question).
For the StarCoder+ model, we allow generation up until the $8192$ context window length, since this model is run locally. In either case, decoding is halted early whenever the stop token \texttt{</EVALUATE>} is produced. All local experiments were executed on a cluster equipped with NVIDIA A100 GPUs.

\textbf{GPT Models}: We use the gpt-3.5-turbo-16k-0613 and gpt-4-0613 checkpoints of the GPT-3.5 \citep{ouyang2022training} and GPT-4 \citep{openai2023gpt4} models, respectively, invoking both models via the OpenAI API. 

\textbf{StarCoder+}: StarCoder+ ($15.5$B)\footnote{\url{https://huggingface.co/bigcode/starcoderplus}} is a version of StarCoderBase \citep{li2023starcoder} which has been finetuned on 600B tokens from a combination of (1) Falcon RefinedWeb \citep{refinedweb} (filtered version of CommonCrawl), (2) The Stack v1.2 \citep{kocetkov2022stack}, and (3) a Wikipedia dataset. Its base model, StarCoderBase, is an open-access model with a GPT-2 architecture using multi-query attention \citep{shazeer2019fast} and fill-in-the-middle objective \citep{bavarian2022efficient}. StarCoderBase has a $8192$ context window and is trained on $1T$ code tokens of permissively licensed text from GitHub across $80$ programming languages \citep{li2023starcoder}. We use StarCoder+ instead of StarCoderBase because it is finetuned on natural language, which should improve the performance on our task. We run StarCoder+ with \texttt{bf16} precision to reduce its memory footprint.