When we released \benchmark{} in January 2024, coding benchmarks primarily measured the
ability to write code. The release of our benchmark introduced
the community to the idea of measuring code reasoning. In the two years since,
the benchmark has shaped the field in three different ways. First, it inspired a family of
code reasoning benchmarks. Second, it became a testbed for many execution-aware training and
fine-tuning advancements. Finally, many frontier model developers adopted it as a standard
evaluation.

\subsection{Inspiring a Family of Code Reasoning Benchmarks}
\benchmark{} was the first benchmark to isolate code reasoning as an independent
capability of interest through input and output prediction. Despite being simple and
easy to run, it revealed many surprising insights about the behavior of language models.
Soon, a wave of benchmarks embraced the idea of code reasoning and extended CRUXEval along
many different axes. CRUXEval-X \citep{xu2024cruxevalx} translated the benchmark into
nineteen programming languages. REval \citep{chen2024reval} increased the granularity of the
prediction target by asking the model to reason about intermediate variable states, coverage, and
execution paths. Others broadened the tested behaviors and semantics:
ThrowBench \citep{prenner2025throwbench} targeted runtime exceptions, CodeMind
\citep{liu2024codemind} added reasoning over a model's own generated code and
specifications, and EquiBench \citep{wei2025equibench} and SeqCoBench
\citep{maveli2025seqcobench} probed functional equivalence between programs.
CodeCrash \citep{lam2025codecrash} tested an adversarial version of \benchmark{} through misleading
perturbations of the original programs. Recently, the idea of code reasoning has also been scaled
beyond simple Python programs to real-world functions \citep{liu2025re2bench}, larger-scale
repositories \citep{li2026reporeason}, and more complicated control flows including parallelism \citep{beger2025coconut}.

\subsection{Explicitly Improving Code Execution}
Since becoming a measurable target, code execution has frequently become
part of the training data for coding models. Several works directly explored
training on execution traces. SemCoder \citep{ding2024semcoder} trained on ``monologue
reasoning'' with execution traces, showing that approximating an interpreter in
natural language improved both \benchmark{} and code generation. Similarly, CodeI/O \citep{li2025codeio}
trained on a large dataset of input-output pairs, showing that it improved not only code
reasoning but also general reasoning across math, logic, and commonsense tasks.
Execution tuning \citep{armengolestape2025execution} trained directly on
execution traces of different granularities, and TracePile \citep{chen2025tracepile}
collected a large-scale dataset of 2.6 million execution trace rationales.
More recently, Meta's Code World Model \citep{faircodegen2025cwm} included execution traces
in its pre-training data to build an internal world model of program semantics.

Since the first wave of works training on execution traces, more creative execution-aware training strategies
have emerged. CodeBoost \citep{wang2025codeboost} incorporated both input and output
prediction on raw code snippets into the reward function of a reinforcement learning loop.
\citet{maimon2026selfexecution} trained models to simulate their own generated
code, improving both \benchmark{} and downstream competitive programming.
Absolute Zero \citep{zhao2025absolutezero} used a single model to self-generate \benchmark{}-style
tasks to train reasoning capabilities with no external data.

\subsection{Adoption by Frontier Model Developers}
Finally, \benchmark{} has been included as a standard evaluation in the release reports of
many frontier models. Just two months after CRUXEval's release, it was included in the
release report of StarCoder2 \citep{lozhkov2024starcoder}. Since then, it has been
included in many other models, such as IBM's Granite Code
\citep{mishra2024granite}, Mistral's Codestral \citep{mistral2024codestral},
01.AI's Yi-Coder \citep{yi2024yicoder}, ByteDance's Seed-Coder
\citep{bytedance2025seedcoder}, Ant Group's Ling-Coder
\citep{codefuse2025lingcoderlite}, Moonshot's Kimi K2
\citep{moonshotai2025kimik2}, Meta's Code World Model
\citep{faircodegen2025cwm}, and successive generations of DeepSeek
\citep{deepseekai2024deepseekv2, deepseekai2024deepseekcoderv2,
deepseekai2024deepseekv3} and Qwen \citep{qwen2024qwen25coder,
qwen2025qwen3technical} models.

Beyond simply reporting scores, a few of these release reports directly discuss
how they optimized for \benchmark{} performance, showing that the benchmark played a role
in shaping model development. For example, Qwen2.5-Coder explicitly mentioned that prioritizing executable
code during data cleaning improved their \benchmark{} performance \citep{qwen2024qwen25coder}.
Ling-Coder-Lite also generated synthetic training data in \benchmark{}'s execution prediction style, which
they said improved their score on the benchmark \citep{codefuse2025lingcoderlite}. Finally, Yi-Coder advertised
being the first open code model to surpass 50\% accuracy on \benchmarko{}
\citep{yi2024yicoder}, showing their attention to the benchmark.
