\chapter{Looking Beyond the Score Reveals Hidden Nuances}
\label{chap:beyond-score}

In Chapter~\ref{chap:right-metrics}, we saw the importance of defining metrics
that capture the properties we actually care about while controlling for confounding biases.
In this chapter, we focus
on the importance of looking beyond the score and analyzing the underlying success and failure 
modes of models. We present four examples showing how deeper analysis can reveal hidden nuances 
in the behavior of language models and advance our scientific understanding.

This chapter first discusses insights and analyses derived from three of the benchmarks 
introduced in the previous chapter. In Section~\ref{sec:cruxeval-beyond-score}, we dive deeper into 
the results from CRUXEval. For example, by comparing scores on CRUXEval with code generation benchmarks like HumanEval,
we discover that a cluster of distilled models (e.g. WizardCoder) perform much better on HumanEval compared
to CRUXEval than their base models, suggesting distillation may not fully transfer general programming abilities.
In Section~\ref{sec:counterfeit-beyond-score}, we analyze whether counterfeit samples have different 
properties depending on the difficulty of the problem or the strength of the model generating the samples. 
In Section~\ref{sec:livecodebench-beyond-score}, we return to the LiveCodeBench paper to analyze
overfitting to HumanEval, the role of post-training, and the gap between open and closed models.
Finally, in Section~\ref{sec:linc-beyond-score}, we introduce LINC, a neuro-symbolic approach to performing
logical reasoning. By analyzing the behavior of LINC compared to traditional chain-of-thought,
we discover that LINC and chain-of-thought have distinct failure modes, suggesting opportunities to
synthesize the two approaches. 


\section{CRUXEval: Generalization, Chain of Thought, Fine-Tuning, and Qualitative Failures}
\label{sec:cruxeval-beyond-score}
\input{paper-includes/cruxeval/beyond-score}

\section{Counterfeit Conundrum: The Nature of Counterfeit Samples}
\label{sec:counterfeit-beyond-score}
\sourcepaper{\path{resources/Counterfeit_Conundrum.zip}. Local source lists Alex Gu as first author, with an equal-contribution author-order note.}{Final thesis prose should be rewritten in dissertation style. If importing close paper phrasing during drafting, wrap it in \texttt{\string\needsrewrite\{...\}} until Alex rewrites it.}
\input{paper-includes/counterfeit/beyond-score}

\section{LiveCodeBench: Overfitting, Post-Training, and Open vs. Closed Models}
\label{sec:livecodebench-beyond-score}
\input{paper-includes/livecodebench/beyond-score}

\section{LINC: A Neurosymbolic Approach for Logical Reasoning by Combining Language Models with First-Order Logic Provers}
\label{sec:linc-beyond-score}
\sourcepaper{\path{resources/_EMNLP_2023__LINC_Logical_Reasoning.zip}. Local source lists equal contribution with randomized author order; Alex Gu can be treated as first author for thesis framing.}{Because parts of the LINC paper prose were written by coauthors, any source-derived LINC text should still be wrapped in \texttt{\string\needsrewrite\{...\}} until Alex rewrites it in his own words.}

\input{paper-includes/linc/beyond-score}

\section{Discussion}
The four examples in this chapter highlight how evaluations can go beyond ranking models and
reveal deeper insights about model behavior. In each case, analyzing the underlying behavior of
models revealed nuances hidden by aggregate scores.
Through CRUXEval, we learned that distillation may not fully lead to transfer between
different programming abilities and saw the potential and limitations of chain-of-thought
prompting and fine-tuning.
In the Counterfeit Conundrum, we saw several counterintuitive trends about the
properties of counterfeit samples and observed that counterfeits are a relatively general phenomenon and
not just a failure mode of weaker models or harder problems. 
In LiveCodeBench, we saw how post-trained models seemed to be overfit to HumanEval, suggesting 
insufficiencies in the diversity of post-training data distributions.
Finally, through LINC, simply looking at scores would show that LINC generally outperformed chain-of-thought
prompting but would miss the nuance that the two approaches were actually complementary and failed
in distinct ways. 
By analyzing the underlying successes and failures in each example, we gained a better high-level understanding
of where methodologies shine and fall short, identifying opportunities for future scientific inquiry.
