\chapter{Conclusion}
\label{chap:conclusion}

I began this thesis with a simple principle: an evaluation is a lens through which we view the
capabilities of a model. As models progressed rapidly, so did the experiments we ran and the
insights that we drew from them. The three lessons in this thesis are not independent guidelines. 
Together, they describe the role I believe evaluations should play in the lifecycle of scientific development. 

First, it is important to choose the right metrics
without bias, because a benchmark that measures the wrong property or contains confounding artifacts may
support a misleading conclusion. Second, we must look beyond the score, because many of the most valuable
scientific insights live beyond surface-level aggregate metrics. Finally, we must design evaluations with the
future in mind, because what we measure shapes what we build.

This final lesson, however, is a double-edged sword. We must keep Goodhart's law in mind:
while a well-designed benchmark can steer an entire community towards a worthwhile capability, when a
measure becomes a target, it ceases to be a good measure. By choosing the right measurement, accounting for
confounding factors, and analyzing beyond the headline metric, we can at least begin to mitigate this problem.

It is worth reflecting on just how quickly the target has changed over the last few years.
When we built CRUXEval, GPT-4 was still failing to solve a third of HumanEval, and it was an open question whether
language models could reason about the code they produced at all. Just a few years later, they are now autonomously solving
competitive programming problems, writing software from scratch, and reasoning about research-level mathematics.
This fast progress shaped the arc of the evaluations in this thesis: going from simulating simple execution behavior to solving
competitive programming problems, proving challenging inequalities, vibe coding applications from scratch, writing iterative
software engineering specifications, and beyond. We continuously adapt the metrics and analysis we choose to the progress in the field.

Indeed, several of the evaluations in this thesis have already become much easier for frontier models. Today's top
models achieve near-perfect scores on evaluations like \benchmark{} and have made substantial progress on
LiveCodeBench. However, their influence has outlasted their original difficulty.
The idea of evaluating code execution and reasoning has been carried forward by a wave of follow-up benchmarks and training methods.
The idea of \textit{live} and contamination-free benchmarks has become a standard way to measure language models as they continue to improve.
A benchmark's value lies not just in how long it resists saturation, but in the lessons and insights it leaves behind.

As models become more capable and autonomous, their inner workings only become less and less transparent.
Therefore, building fresh evaluations that serve as reliable lenses into these systems will be increasingly crucial.
While the benchmarks in this thesis have already begun to saturate, the discipline of building them will remain intact.
