\chapter{Introduction}
\label{chap:introduction}

In June 2020, OpenAI released GPT-3 \citep{brown2020language}, a language model that
felt strong for its time but was still unreliable for many knowledge-based tasks.
In the two years that followed, models showed progress through systems like Codex, Copilot, 
and InstructGPT, but they remained brittle for programming, mathematics, and deeper
reasoning tasks. In November 2022, however, a breakthrough seemed to occur. ChatGPT was 
released, and for the first time an AI system felt like a generally usable assistant. 
Just four months later, in March 2023, GPT-4 was made available, bringing a large jump 
in reliability and performance.

Suddenly, the pace of progress made it seem like models could do anything, faster than our benchmarks
could keep up. By the time GPT-4 was released, the community had built a
substantial evaluation ecosystem: MMLU \citep{hendrycks2020measuring} tested broad academic knowledge, 
BIG-Bench \citep{srivastava2022beyond} and BIG-Bench Hard \citep{suzgun2022challenging} collected diverse 
challenging tasks, and HELM \citep{liang2022holistic} organized many evaluations into a more systematic comparison 
framework. Other benchmarks targeted narrower capabilities that had long been difficult for language models, from 
commonsense and science reasoning in WinoGrande \citep{sakaguchi2021winogrande}, HellaSwag \citep{zellers2019hellaswag}, 
and ARC Challenge \citep{clark2018think}, to grade-school mathematical reasoning in GSM8K \citep{cobbe2021training}, 
to code generation in HumanEval \citep{chen2021evaluating}, MBPP \citep{mbpp}, and APPS \citep{hendrycks2021measuring}. 
Many of these evaluations had been designed as long-term challenges and stood for years, yet were now starting
to saturate. On MMLU, GPT-4 reached 86\%, approaching the $\sim$90\% accuracy of human domain experts and nearly
doubling the original GPT-3's 44\%. It exceeded 85\% on GSM8K and WinoGrande and surpassed 95\% on HellaSwag and
ARC Challenge. The same pattern held for code: on HumanEval, a popular coding benchmark, GPT-4 scored 67\%, up
from GPT-3.5's 48\% and the original Codex's 29\%.

At the same time, the models seemed to be able to do much more than what the benchmarks had sought
to measure. One example of this was the ``Sparks of AGI'' report \citep{bubeck2023sparks}, which
set aside standardized benchmark-based evaluation in favor of impressive qualitative demonstrations 
of GPT-4: drawing a TikZ unicorn, writing a poem-based proof that there are infinitely many primes, 
coding a functioning game from scratch, and displaying theory of mind. For a while thereafter, the best 
way to understand these models was not to rely on academic benchmarks, but rather to browse community 
demonstrations of what could be accomplished.

The perception in the academic community therefore shifted. Previously, each individual research 
community (NLP, coding, vision) developed models more or less with open recipes, guided by benchmarks 
that served as a north star for progress. GPT-4 broke this status quo. It was the strongest system across 
all of these domains at once, yet almost nobody outside of OpenAI understood how it had been trained. 
With the strongest systems now closed and hard to understand from the outside, what we needed from an
evaluation was itself changing: beyond serving as a guide to scientific progress, evaluations increasingly
had to help us make sense of behaviors and quirks we could only observe externally. More broadly, it was a
ripe time to rethink the role of evaluations.

Ultimately, an evaluation needs to help model builders and scientists understand the
capabilities and limitations of a model. The challenge is to design evaluations that
identify meaningful shortcomings, advance our scientific understanding of the current
state of the art, and suggest avenues for moving forward as a field. In this
thesis, I will share three lessons from my experience designing evaluations for language models
in the several years that followed GPT-4, particularly for coding, mathematics, and logical reasoning.


\section{Lesson 1: Focusing on the Right Metrics Without Bias Is Crucial}

The first lesson is that focusing on the right metrics without bias is crucial. In science,
almost anything can be measured, but not everything is meaningful. The first question we face when designing any
evaluation is what we should measure. The second is whether the resulting measurement is biased by
confounding factors in the data or evaluation setup.

In late 2023, language models like GPT-4 were just starting to become proficient at writing Python code.
When humans learn programming, they not only learn to write code, but also gain the ability to read code, understand it,
simulate its execution behavior, debug it, and develop a suite of other programming-related skills. At the time, there was little knowledge about 
whether language models learned to do programming in the same ways that humans did. Especially because the coding data that language models were
trained on primarily consisted of GitHub code, it was unclear whether these models could generalize to capabilities beyond code generation.
Motivated by this, we designed \benchmark, a benchmark asking language models to simulate the forward and reverse execution of very simple
Python programs consisting of 3-13 lines with no complex control flow or calculations.

When we released \benchmark, many people in the community were skeptical about the design and utility of such a benchmark.
After all, a simple way to ace the forward execution subset of \benchmark{} is to simply run the code in a Python interpreter.
However, we still saw value in this benchmark for several reasons. First, as is evident today, language models that can effectively 
program should have a good understanding and mental model of code. This may have been overlooked when simple single-turn benchmarks
like HumanEval were the standard for coding evaluations, because real-world programming requires a much wider set of skills beyond
single-shot code generation. Second, the benchmark allows us to scientifically understand new phenomena of coding language models.
For example, by showing that language models did have some ability to perform both forward and reverse execution, we saw early signs that
language models did have some degree of generalization despite just training on raw code tokens.

Beyond choosing the right metric, we must also account for potential biases in our evaluation design before arriving at conclusions.
As an example, in IneqMath, we introduced a benchmark for evaluating the ability of language models to solve inequality problems. 
When we measured the accuracy of Grok 3 mini, we saw that it was able to get the answer correct $72\%$ of the time, indicating a high level of proficiency at solving inequalities.
However, upon closer examination, when changing the metric from evaluating at the final answer to evaluating the entire reasoning trace and solution, we found that the
accuracy dropped to $6\%$, a stark difference that leads to an entirely different conclusion. Arriving at the first conclusion misses the fact that models
may be able to correctly guess the final answer without providing a correct proof.

Another example showing the importance of accounting for confounding factors is our LiveCodeBench evaluation.
Prior to LiveCodeBench, coding evaluations primarily relied on static benchmarks like HumanEval, fixed datasets that were released once and 
never updated. However, this fails to account for data contamination: language models are often trained on data from the internet, which 
could potentially include these benchmarks. While many models are trained with best-effort decontamination processes, it is not possible to 
completely remove all traces of these benchmarks from the training set. Using time-controlled evaluations, we eliminate this bias. 
As evidence of this, DeepSeek-Instruct-33B has a published data cutoff date of September 2023. When evaluated on problems between May and August 2023, it had an average 
accuracy of around 50\%. However, when evaluated on problems between September 2023 and February 2024, the average accuracy dropped to around 20\%, significantly lower.
Similarly, the accuracy of DeepSeek-Base-33B (same cutoff date) plummeted from a pass@1 of 60\% on May 2023 LeetCode problems to around 0\% on September 2023 problems.

These examples preview the main idea of Chapter~\ref{chap:right-metrics}, but they are not exhaustive.
The chapter will also discuss how iterative code-quality metrics in SlopCodeBench and counterfeit-focused
metrics in the Counterfeit Conundrum expose failures that standard accuracy would obscure. Together, these
case studies show that the right metric and the right controls can change the conclusion one draws from an evaluation.
Once the metric exposes a gap, the next step is to analyze the behavior that produced it.
Through in-depth analysis, we can often discover phenomena that are not obvious from the final
scores alone, leading to the next lesson.

\section{Lesson 2: Looking Beyond the Score Reveals Hidden Nuances}

Often, the most interesting scientific insights only appear after looking beyond surface-level metrics
with deeper analysis. In \benchmark{}, the execution metric already revealed that models could perform some execution;
deeper analysis revealed more about how different coding capabilities relate to one another. The most striking finding
came when we looked at the relationship between HumanEval scores and \benchmark{} scores across a suite of models. While the
forward execution accuracy was highly correlated with the reverse execution accuracy, the HumanEval score showed a different
trend. For base models, HumanEval and \benchmark{} performance were correlated.
However, distilled models (WizardCoder, Phind, Phi) broke this trend, significantly beating their base models at HumanEval but
not \benchmark. This reveals a key limit of fine-tuning: while distillation on NL-to-code samples can greatly boost performance on
code generation, it does not necessarily improve a model's underlying understanding of code.

Another illustration of this is when we developed LINC, a neurosymbolic method for logical reasoning.
The task is that, given a set of natural language premises and a conclusion, one must determine if the conclusion
is true, false, or unknown based on the premises. The standard method of doing this was via chain-of-thought (CoT)
prompting, having a language model reason through the premises to make a deduction about the validity of the conclusion.
Our method, LINC, first formalized everything into a first-order logic program and then executed a first-order logic prover
to determine the answer. In most evaluations, LINC outperformed CoT, which might lead one to believe it is
a superior method for performing logical reasoning. However, in one case, on the FOLIO benchmark 
using GPT-4 as the base model, this was not the case: LINC only scored $73\%$ vs. CoT's $75\%$. 

After this result, we conducted a series of analyses to understand why LINC had underperformed in this
specific case. First, we did a qualitative analysis of the errors of both approaches, and discovered they were very 
different. Errors in LINC occurred when the models failed to properly capture implicit and explicit information that was 
in the natural language statements (failure modes impossible in CoT), while errors in CoT occurred from making incorrect natural language logical deductions
and failing to find the right reasoning chain (failure modes impossible in LINC). Upon further quantitative analysis, we found
that LINC had worse recall but better precision on True/False predictions, and that LINC and CoT mispredicted on different examples.
These nuances were not apparent from looking at the performance scores and only came to light after
this deeper analysis.

Chapter~\ref{chap:beyond-score} expands this lesson beyond these two examples. We return to the
Counterfeit Conundrum to understand what kinds of wrong programs models find confusing, and to
LiveCodeBench to analyze overfitting, post-training, and differences between open and closed models.
The common thread is that the aggregate score is a starting point: the scientific value often comes
from analyzing the structure of the successes and failures behind it.

Designing the right metric and carrying out the right analysis is only the first step.
Once an evaluation reveals these nuances, it naturally raises further questions about what
capabilities are missing, which failure modes matter most, and what kinds of systems should be
built next. In this way, evaluations not only describe the current state of models, but also
influence the trajectory of future research. This is the final lesson: what we measure
shapes what we build.

\section{Lesson 3: What We Measure Shapes What We Build}

Evaluations are most useful when they are forward-looking. Once a benchmark is released, the
capabilities and failure modes it makes visible often become focal points for the relevant
scientific community, influencing both what researchers measure and what model-building labs
optimize for. This does not mean that a benchmark must remain difficult forever. Often, its
long-term value is that it names a capability, makes a failure mode concrete, and gives the field
a shared target to improve.

\benchmark{} is one example. By measuring code execution and input-output reasoning directly, it
helped make execution reasoning a target for both evaluation and training. Several later benchmark
efforts explicitly built on this target: CRUXEval-X \citep{xu2024cruxevalx} extends execution
evaluation to 19 programming languages, REval \citep{chen2024reval} evaluates runtime state
prediction in addition to input-output prediction, and CodeMind \citep{liu2024codemind} introduces
additional code-reasoning tasks such as specification reasoning and dynamic semantics reasoning.
The same idea also influenced diagnostic work. For example, in the Counterfeit Conundrum
\citep{gu2024counterfeit}, we used execution as a probe to show that language models often simulate
subtly incorrect programs as if they were correct, and SeqCoBench \citep{maveli2025seqcobench}
tested whether code LLMs recognize functionally equivalent code under semantics-preserving
transformations.

The influence was not limited to evaluation. Follow-up training papers used execution traces and
input-output prediction as supervision for stronger code reasoning models: SemCoder
\citep{ding2024semcoder} teaches models to reason about program semantics, CodeI/O
\citep{li2025codei} trains on input-output prediction, and Self-Execution Simulation
\citep{maimon2026selfexecution} combines execution-trace fine-tuning with reinforcement learning for
self-repair. \benchmark{} also became a standard diagnostic benchmark reported by many new coding
model releases, including DeepSeek-Coder V2 \citep{deepseekai2024deepseekcoderv2}, Qwen 2.5
\citep{qwen2024qwen25coder}, Qwen 3 \citep{qwen2025qwen3technical}, Kimi K2
\citep{moonshotai2025kimik2}, StarCoder 2 \citep{starcoder2}, and Code World Model
\citep{faircodegen2025cwm}. In this sense, making execution reasoning measurable helped turn it
into a capability that later systems were trained for, reported on, and optimized around.

Chapter~\ref{chap:measure-shapes} develops this lesson through a broader set of examples rather
than only through \benchmark{}. LiveCodeBench shows how contamination-aware, time-indexed evaluation
became a standard way to measure coding and reasoning models. VibeCodeBench shifts the target from
isolated coding tasks to full web application development. LeanDojo illustrates how an evaluation can
also become infrastructure for an entire research area. ProofOptimizer shows how measuring proof
length turns proof simplification into a concrete optimization objective. Together, these examples
support the final lesson of the thesis: what we measure shapes what we build.

\section{Works Covered in This Thesis}

The technical chapters of this thesis synthesize a collection of research projects on evaluating
language models for coding, mathematics, and reasoning. Chapter~\ref{chap:right-metrics} introduces \benchmark{}
\citep{gu2024cruxeval}, SlopCodeBench \citep{orlanski2026slopcodebench}, IneqMath \citep{lu2025ineqmath},
the Counterfeit Conundrum \citep{gu2024counterfeit}, and LiveCodeBench \citep{jain2024livecodebench}.
Chapter~\ref{chap:beyond-score} returns to several of these works and also introduces LINC \citep{olausson2023linc}.
Chapter~\ref{chap:measure-shapes} additionally introduces VibeCodeBench \citep{vibe-code-bench}, LeanDojo \citep{yang2023leandojo}, and
ProofOptimizer \citep{gu2026proofoptimizer}. Chapter~\ref{chap:position-paper} closes with the forward-looking report
\textit{Future Research and Remaining Challenges Toward AI for Software Engineering} \citep{gu2025positionaiswe}.

These projects were not done alone. Many of the works discussed in this thesis were collaborations,
and the full author lists in the cited papers reflect the people who helped shape the ideas, build the
systems, run the experiments, and write the original papers. Throughout the thesis, ``we'' refers to the research teams behind the corresponding
works, and I am grateful to all of these collaborators for their contributions to all the content in this thesis.
