\chapter{Focusing on the Right Metrics Without Bias Is Crucial}
\label{chap:right-metrics}

Evaluation metrics play a pivotal role in shaping a scientific community's collective intuition.
They determine which aspects of model behavior become visible and how those behaviors are
interpreted. Designing new metrics can surface surprising capabilities and shortcomings, while
careful scrutiny of existing metrics can reveal confounding factors that distort evaluation
outcomes and lead to misleading conclusions about model performance. It is therefore important to
both create fresh metrics and remove sources of bias from existing ones, so that benchmark
results faithfully reflect the abilities they are intended to measure.
As mentioned in Section~\ref{sec:intro-right-metrics-without-bias}, I use \textit{bias}
in this thesis to refer to a systematic confounding factor that makes a score reflect
something other than the intended measurement, as opposed to statistical or social bias.

In the first three sections, we introduce metrics that capture desired
properties and capabilities missed by existing benchmarks. Section~\ref{sec:cruxeval-code-reasoning-understanding-execution-evaluation}
introduces \benchmark, a benchmark that measures code reasoning, understanding, and execution.
Through this work, we provided evidence that coding language models had just started to develop
the ability to execute simple Python programs while showing that top models at the time still
had significant gaps. Section~\ref{sec:slopcodebench-right-metrics} introduces SlopCodeBench,
where we measured two metrics related to code quality: verbosity and structural erosion. By analyzing
these metrics throughout iterative coding trajectories, we discovered that even if models can write correct code,
the quality of that code degrades over time, which has implications for the maintainability of code generated by language models.
In Section~\ref{sec:ineqmath-right-metrics}, we discuss IneqMath, a benchmark of inequality proving problems.
From measuring both final-answer correctness and step-wise reasoning soundness, we revealed that LLMs frequently
arrive at the correct answer but hallucinate unsound reasoning steps.

In the final two sections, we turn our attention toward faithful measurement by isolating
confounding factors that change how standard metrics should be interpreted. Section~\ref{sec:counterfeit-conundrum} introduces the
\textit{counterfeit conundrum}, where we found that filtering out trivial incorrect samples in the
benchmark can greatly change how we perceive accuracy on tasks like correctness checking, code execution,
and program repair. Finally, Section~\ref{sec:livecodebench-right-metrics} discusses LiveCodeBench, where we
measured accuracy across time on a suite of code-related tasks like code generation, revealing that accounting for
contamination effects can drastically affect how models compare against each other.

\section{CRUXEval: A Benchmark for Code Reasoning, Understanding and Execution}
\label{sec:cruxeval-code-reasoning-understanding-execution-evaluation}
\input{paper-includes/cruxeval/right-metrics}

\section[SlopCodeBench: Benchmarking How Coding Agents Degrade Over Long-Horizon Iterative Tasks]{SlopCodeBench: Benchmarking How Coding\\ Agents Degrade Over Long-Horizon Iterative Tasks}
\label{sec:slopcodebench-right-metrics}
\sourcepaper{\path{resources/latex_slop_code.zip}. Alex contributed to SlopCodeBench but is not first author; local source lists Gabriel Orlanski before Alex Gu.}{Non-first-author source: keep any source-derived prose marked in blue and rewrite/attribute it before final submission.}

\input{paper-includes/slopcodebench/right-metrics}

\section{IneqMath: Solving Inequality Proofs with Large Language Models}
\label{sec:ineqmath-right-metrics}

\input{paper-includes/ineqmath/right-metrics}

\section{The Counterfeit Conundrum: Can Code Language Models Grasp the Nuances of Their Incorrect Generations?}
\label{sec:counterfeit-conundrum}
\input{paper-includes/counterfeit/right-metrics}

\section{LiveCodeBench: Holistic and Contamination Free Evaluation of Large Language Models for Code}
\label{sec:livecodebench-right-metrics}
\input{paper-includes/livecodebench/right-metrics}

\section{Conclusion}

An important aspect to highlight is that the metrics we focused on needed to change over time to adapt to the
changing landscape of frontier model development. When we developed CRUXEval, the standard coding evaluations were
single-function static benchmarks. The community had not thought deeply about capabilities beyond code generation, and
CRUXEval helped signal the importance of broader capabilities such as code execution. As discussed later in
Section~\ref{sec:cruxeval-measure-shapes}, this mattered not only because it revealed a missing ability at the time,
but also because it helped make execution behavior a target for later benchmarks and training recipes. In particular,
CRUXEval helped motivate work on training models with execution traces, one of the ways code LLMs improved in the years
that followed.

When we later saw that there were several biases distorting how we interpreted accuracy, we turned towards eliminating biases
through works like the Counterfeit Conundrum and LiveCodeBench. The contamination-free evaluation methodology from
LiveCodeBench quickly gained wide adoption, as discussed in Section~\ref{sec:livecodebench-measure-shapes}.
Accurate model comparisons became more important as capabilities scaled up.
In the following year, models quickly improved and were increasingly adopted by software engineers, making it clear that single-function static
benchmarks were no longer reflective of real-world use. This shift motivated SlopCodeBench, where we focused on
iterative evaluation and code-quality metrics that better align with long-horizon developer workflows and the priorities
of software engineers. Overall, the metrics that we designed evolved alongisde the various shifts at the frontier.

In each of the five preceding sections, we saw how important it is to focus on the right metric and eliminate confounding
sources of bias. By deciding on a metric, we control what we can learn. The first three evaluations showed the presence
and absence of particular abilities: models were beginning to execute simple code snippets, could not maintain code quality
over long iterative trajectories, and were often better at finding the right answer to inequalities problems than at
producing sound stepwise proofs. The final two evaluations highlighted biases that changed how we perceived the results of
our measurements. The next chapter looks at subtle model behaviors exposed by deeper analysis into these metrics.
