\input{paper-includes/livecodebench/macros-import}
Around the time of \livecodebench{}'s release, models were being trained
on larger and larger crawls of the internet, making static benchmarks increasingly vulnerable
to contamination. It was the first benchmark to challenge the status quo of static coding
benchmarks, instead focusing on \emph{live}ness, continuously collecting new problems and
updating the evaluation set every few months. First, this design inspired a wave of
contamination-aware benchmarks across code, math, and general reasoning, helping to shift
the community's mindset to treat contamination as a first-class concern.
Second, with over $1500$ citations, \livecodebench{} has been adopted as a standard evaluation by
researchers and frontier model developers alike, utilizing appropriate time windows
to avoid contamination.

\subsection{Contamination-Aware Benchmarks}
\livecodebench{} popularized the recipe of using problems from a scrolling
time window drawn from a continuously updated source of new tasks. LiveBench
\citep{white2024livebench} extended \livecodebench{} beyond code to a larger suite of tasks
including math, reasoning, language, and data analysis. The recipe also propagated into adjacent domains: LiveMathBench
\citep{liu2024livemathbench} applied the same idea to competition
mathematics, assembling fresh problem sets from annual contests such as the AMC and the
Putnam. LiveAoPSBench \citep{mahdavi2025liveaopsbench} timestamped olympiad-level
problems mined from the Art of Problem Solving forum, similarly observing a steady
performance decline on more recent problems and showing a similar phenomenon in the math domain.
LiveMathematicianBench \citep{he2026livemathematicianbench} pushed the idea to
research-level mathematics, drawing problems from arXiv papers posted after model
training cutoffs.
LiveCodeBench Pro \citep{zheng2025livecodebenchpro} extended the benchmark's idea to harder olympiad-level
problems, and in software engineering, SWE-bench-Live
\citep{zhang2025swebenchlive} applied the same monthly-update discipline to
real GitHub issues created after a fixed date. The idea has even been applied to multimodal
evaluation: LiveXiv \citep{shabtay2024livexiv} built a continuously refreshed
visual-question-answering benchmark from newly posted arXiv manuscripts.

Beyond the specific benchmarks it inspired, \livecodebench{} helped shift how
the community reasons about contamination. Contamination itself was not a new
concern: prior studies had developed methods to detect benchmark data in
training corpora \citep{golchin2023time}, found that performance on coding
problems correlates with their prevalence on GitHub \citep{roberts2024to,
riddell2024quantifying}, and caught coding benchmarks leaking into training
data shortly after release \citep{matton2024leakage}. However, these works
treated contamination as something to detect after the fact. Instead,
\livecodebench{} treated contamination as a first-class design concern.
By showing a scrolling benchmark with time-adjustable windows, it directly
exposed sharp performance drops on newer problems. As evident by the
number of live benchmarks following \livecodebench{}, contamination-free benchmarks
have become a standard expectation.

\subsection{Adoption by Frontier Model Developers}
Finally, \livecodebench{} became part of the standard evaluation suite for
frontier model releases, particularly among open-weight and
reasoning-oriented systems. It appeared in the release reports of successive
DeepSeek generations \citep{deepseekai2024deepseekv3,
deepseekai2025deepseekr1}, the Qwen2.5-Coder and Qwen3 reports
\citep{qwen2024qwen25coder, qwen2025qwen3technical}, Google's Gemini 2.5
\citep{comanici2025gemini}, Moonshot's Kimi K2 \citep{moonshotai2025kimik2},
ByteDance's Seed-Coder \citep{bytedance2025seedcoder}, Zhipu's GLM-4.5
\citep{zeng2025glm45}, MiniMax-M1 \citep{minimax2025m1}, Mistral's Magistral
\citep{mistral2025magistral}, and NVIDIA's Llama-Nemotron
\citep{bercovich2025llamanemotron}, among others.

In addition, many of these reports referenced \livecodebench{}'s contamination methodology.
The DeepSeek-V3 report \citep{deepseekai2024deepseekv3} restricted evaluation to a
recent window (problems from August to November 2024) precisely to avoid
leakage. Similarly, the Qwen2.5-Coder report \citep{qwen2024qwen25coder}
evaluated only on the most recent months, describing them as questions that ``could
not have leaked into the training set'' and thus reflect
out-of-distribution ability. Gemini 2.5 \citep{comanici2025gemini} reported
\livecodebench{} with its exact problem date range (October 2024 through
January 2025) stated alongside the score.
