\section{Iterative Evaluation of Coding Agents}\label{sec:results}

This section reports empirical results on \scb{}. The evaluation targets
behaviors that arise once an agent edits its own code across multiple
checkpoints, a regime that one-shot benchmarks do not exercise. Four
research questions structure the experiments, each stated with its
supporting result.

\noindent\textbf{RQ1: Can agents iteratively extend their own design decisions?} No configuration is able to produce a solution that passes every checkpoint for any problem. Each checkpoint's cost rises without corresponding improvements to correctness.

\noindent\textbf{RQ2: Does iteration cause agent code to degrade?} Structural erosion increases in \erosionPctIncreasing\% of trajectories while verbosity increases in \vfTrajIncPct\% of trajectories.

\noindent\textbf{RQ3: How does agent degradation compare to repository histories?} Compared to \avhPanelN{} Python repositories, agents produce code that is \avhAllErosionRatio{}$\times$ more structurally eroded and \avhAllVerbRatio{}$\times$ more verbose. Agents additionally exhibit significantly higher degradation, especially compared to pre-2024 commits.

\noindent\textbf{RQ4: Can better prompts stop this behavior?} Better prompts can reduce initial structural erosion by up to \psRqInitialEroReductionMax\% and verbosity by up to \psRqInitialVerbReductionMax\%, but do not stop iterative degradation. The average quality velocity remains \psRqQualityVelocityPp{} percentage points per checkpoint. Different prompting methods additionally increase the cost per checkpoint by \psRqCostIncreaseMean\% on average.


\paragraph{Setup.} 

Each checkpoint runs in a fresh Docker container as a non-root user; only the working directory persists between checkpoints. Frontier models are trained for their provider's harness rather than generalized agent loops, and most developers interact with agents through these CLI tools. We therefore evaluate in native harnesses rather than frameworks such as SWE-agent~\citep{swe-agent}. We evaluate \overallNumModels{} models across six providers in detail. Each run has a two-hour wall-clock limit, no turn or cost cap, and uses the \texttt{just-solve} prompt, which only instructs the agent to solve the problem and keep track of dependencies. Additional setup details are found in \autoref{sec:eval-env-details}.

Problems range from 3 to 8 checkpoints, so raw checkpoint indices are not directly comparable across problems. For aggregation and visualization, we map each trajectory to one of five \emph{progress phases}. The first checkpoint is always \textbf{Start} and the last is always \textbf{Final}. The remaining interior checkpoints are divided into three groups: \textbf{Early}, \textbf{Mid}, and \textbf{Late}. All per-phase statistics in this paper use this.

\input{sections/results/overall}
\input{sections/results/quality_accumulation.tex}
\input{sections/results/agent_vs_human.tex}
\input{sections/results/prompt_strategies.tex}
