\section{Results Continued}\label{sec:human-repo-panel}

\input{figures/performance_table_appendix.tex}


\subsection{Human Panel construction}

The temporal panel spans \avhPanelN{} maintained Python repositories sampled across web frameworks, scientific computing, infrastructure tools, machine-learning libraries, and command-line utilities. Repositories are stratified by GitHub star tier: Hobby ($<100\bigstar$, $n{=}\avhHobbyN$), Niche ($100$--$1\text{k}\bigstar$, $n{=}\avhNicheN$), Established ($1$--$10\text{k}\bigstar$, $n{=}\avhEstabN$), and Major ($\geq10\text{k}\bigstar$, $n{=}\avhMajorN$). For each repository we sample up to 30 source-touching commits across its history, totaling \avhTemporalCkpts{} checkpoints between \avhPanelStartYear{} and \avhPanelEndYear{}. Checkpoints with empty source trees or failed metric runs are dropped before analysis. Each repository's most recent checkpoint serves as its HEAD snapshot for the calibration band reported in \autoref{sec:agent-vs-human}. \autoref{tab:agent-vs-human} reports HEAD verbosity and erosion by tier and against agent checkpoints.


\paragraph{Pre- vs.\ post-ChatGPT temporal check.} The panel straddles 2022, so post-ChatGPT commits could in principle include LLM-assisted contributions. Pre-ChatGPT checkpoints ($n{=}\avhTemporalPreN$) have a median verbosity of \avhTemporalPreVerbMedian{} versus \avhTemporalPostVerbMedian{} for post-ChatGPT checkpoints ($n{=}\avhTemporalPostN$). The shift is real but small relative to the agent gap (mean verbosity \avhAgentVerbMean{}). The same comparison anchored at January 2024 yields medians of \avhAgentEraPreVerbMedian{} versus \avhAgentEraPostVerbMedian{} for verbosity and \avhAgentEraPreErosMedian{} versus \avhAgentEraPostErosMedian{} for erosion.

\paragraph{Within-repository agent-era shift.} A cross-section across the 2024 boundary mixes new repositories in with old ones. Restricting to the \avhAgentEraBothN{} repositories with at least three checkpoints in each era, the median within-repository verbosity shift is $+\avhAgentEraVerbShiftMedian$ and the erosion shift is $+\avhAgentEraErosShiftMedian$. \avhAgentEraVerbHigherPct{}\% of these repositories are more verbose post-2024 and \avhAgentEraErosHigherPct{}\% are more eroded. The signal is consistent with a small uptick in human-authored slop after agent-assisted contributions became common, but it is dwarfed by the agent-side trajectory growth reported in \autoref{sec:agent-vs-human}.

\paragraph{First-to-last growth.} For repositories with at least five sampled commits, \avhTemporalVerbRising{} of \avhTemporalVerbTotal{} ($\avhTemporalVerbRisingPct{}\%$) end with higher verbosity than they started, with median first-to-last growth of \avhTemporalVerbGrowthMedian{}\%. Erosion rises in \avhTemporalErosRising{} of \avhTemporalErosTotal{} ($\avhTemporalErosRisingPct{}\%$) repositories. Per-checkpoint slope is \avhHumanVerbSlopeMedian{} for verbosity and \avhHumanErosSlopeMedian{} for erosion across the human panel, against \avhAgentVerbSlopeMedian{} and \avhAgentErosSlopeMedian{} for agent trajectories.

\paragraph{Notable outliers.} \texttt{huggingface/transformers} (verbosity \avhTransformersVerb{}, erosion \avhTransformersErosion{}) and \texttt{fastapi/fastapi} (\avhFastapiVerb{}, \avhFastapiErosion{}) are the most prominent verbosity outliers in the panel, both above the agent verbosity mean. The high transformers verbosity is driven by per-architecture file duplication (one model implementation per family). At the other end, \texttt{pallets/flask} (\avhFlaskVerb{}, \avhFlaskErosion{}) and \texttt{scikit-learn/scikit-learn} (\avhSklearnVerb{}, \avhSklearnErosion{}) are large, mature projects with low verbosity and middling erosion. \texttt{scipy/scipy} (\avhScipyVerb{}, \avhScipyErosion{}) sits at the high end of human erosion, still well below the agent mean.
