\section{Extended Related Work}\label{sec:appendix-extended-related-work}

\paragraph{Quality degradation under multi-turn coding.} LLM-generated code degrades under repeated modification. Code converges toward structural attractors~\citep{iterative-readability}, quality diverges across agents and runs~\citep{chen2025-patch-quality, santos2025-software-aging, bohr2025-show-and-tell}, and refinement introduces defects that correctness testing does not catch~\citep{scaffold-cegis, diff-fuzzing-refactoring}. Interaction failure modes compound these effects~\citep{interaction-smells, overcorrection-reviewers, more-rounds-more-noise}; agent-generated code has become a practical integration burden~\citep{nakashima2026-agentic-prs, vibe-coding-overhead}.

\paragraph{Code quality metrics.} LLM code has higher cyclomatic complexity per line than human-written code~\citep{dou2024-whats-wrong-llm-code}, though aggregate CC can fall while vulnerabilities rise~\citep{cotroneo2025-human-vs-ai}. \citet{abbassi2025-taxonomy-inefficiencies} extend the code smell taxonomy to LLM outputs, finding redundant steps, duplication, and unnecessary conditionals most prevalent. Code smells are well-studied in software engineering~\citep{fowler1999-refactoring, lacerda2020-code-smells-tertiary}; software aging~\citep{parnas1994-software-aging} and technical debt~\citep{cunningham1992-technical-debt} describe how structural degradation accumulates under modification, with \citet{li2022-architecture-erosion} and \citet{le2021-architectural-decay} documenting this in long-lived systems.

\paragraph{Benchmark landscape.} Beyond the paradigm established by \citet{swe-bench}, the evaluation landscape spans three waves. The first broadens language and domain coverage through extensions~\citep{swe-bench-plus, multi-swe-bench, rust-swe-bench, swe-rebench-v2, swe-polybench}. Instruction-following benchmarks evaluate compliance across conversation turns~\citep{codeif-bench, multicodeif} but assess each response independently. A second wave targets feature-level development from existing repositories~\citep{fea-bench, featurebench}. A third wave builds entire projects from scratch: \citet{commit0} generate libraries from specifications with interactive test feedback, several benchmarks construct full repositories from natural-language requirements~\citep{projdevbench, e2edevbench, nl2repo-bench, swe-agi, projecteval}, and \citet{longcli-bench} spans from-scratch construction through refactoring. Tasks grow larger and harder across all three waves; evaluation in each case is a single artifact assessed against a fixed specification.
