\subsection{Paths Forward} \label{sec:paths}

The challenges above suggest a few paths forward organized around a common theme: code models need feedback-rich training and evaluation, not only more examples of final code. Future systems should be built around execution, code quality, tool use, and semantic understanding across larger software contexts. I focus here on four directions: collecting richer code data, training in executable environments, adapting to specialized codebases, and integrating retrieval and tools into the model workflow.

\subsubsection{Data Collection} \label{sec:d-data}

\begin{aicodeorangebox}
One bottleneck in the development of AI for SWE in the open-source community is the lack of access to fine-grained and high-quality code data. Automated techniques can augment existing programs with symbolic information and generate synthetic data with verifiers, while human-centered datasets can capture signals that are difficult to infer from static GitHub snapshots.
\newline \\
\textit{Challenges addressed: \ref{sec:c-evaluation}, \ref{sec:c-global-understanding}}
\end{aicodeorangebox}

\par\medskip\noindent\textbf{Automatic Data Curation}\phantomsection\label{sec:d-automatic-data}\par\smallskip

\textbf{Augmenting data with program information}: One challenge in enabling LLMs to develop a world model of code is that programs are often treated like text: as tokens with no semantic information. However, modern programming tools allow us to extract rich semantic and structural information about code, such as abstract syntax trees, control-flow graphs, type information, data-flow facts, runtime traces, memory usage, coverage, call stacks, and program invariants. By leveraging these tools, we can augment training datasets with annotations describing how programs execute and how different pieces of code relate to one another. This direction is also reflected in the CRUXEval follow-up work discussed in Sec.~\ref{sec:cruxeval-execution-impact}: several works have begun curating execution traces, input-output pairs, and trace rationales as training data, rather than using execution only as an evaluation signal.

\textbf{High-quality, verifiable synthetic data}: The advantage of code is that it is possible to obtain strong feedback with test cases, program execution engines, type checkers, proof assistants, and other symbolic tools. This makes high-quality synthetic data generation viable, because a large batch of generated data can be filtered using verifiers. For example, to generate code with interesting program invariants, we can sample programs, run an invariant detection engine, and retain only programs with nontrivial invariants. While synthetic data in code has mostly been at the function-level scope, there are no fundamental bottlenecks to expanding to larger scopes. As code is compositional, individual building blocks can be combined to generate synthetic data at the repository-level scope, which can be useful for both training and evaluation.

While the importance of having high-quality data versus high quantities of data is debated, verified data has proven useful. For example, \citet{code-r1} show that simply removing bugs in existing datasets such as TACO \citep{li2023taco} can lead to significant boosts. KodCode \citep{xu2024benchmark} also showed that fine-tuning on verified synthetic data leads to significant improvements.

\par\medskip\noindent\textbf{Human-Centric Data Curation}\phantomsection\label{d:subsec-human-data}\par\smallskip

Many code LMs are trained on datasets such as \textit{the Stack} \citep{kocetkov2022stack, lozhkov2024starcoder}, consisting of trillions of tokens sourced from GitHub. However, training on raw GitHub tokens omits many crucial human signals in the process of software development. For example, companies such as Google rely on internally captured logs of high-quality SWE data, including fine-grained code edits, build outcomes, copy-paste actions, code reviews, reviewer fixes, and change submissions \citep{chandra2024ai}. Similarly, Meta and GitHub Copilot use telemetry with their AI coding assistants to track and leverage signals from AI-generated code \citep{murali2024ai, ziegler2024measuring}. These data sources can reveal not only whether code was accepted, but also how it was revised, tested, reviewed, and maintained over time.

Collecting this data raises privacy and intellectual-property concerns, but community efforts around task-specific datasets, anonymized traces, and executable open-source repositories could still make these signals more broadly available.

\subsubsection{Training}

\par\medskip\noindent\textbf{Environment Design for Code RL}\phantomsection\label{sec:d-rl}\par\smallskip

\begin{aicodeorangebox}
Reinforcement Learning from Verifiable Rewards (RLVR)~\citep{lambert2025tulu3pushingfrontiers} has emerged as a powerful paradigm in math and coding domains where model outputs can be evaluated against a ground-truth outcome, such as exact match or passing a set of unit tests. For code, promising directions include collecting executable codebase environments, sourcing task prompts and rewards from repositories, and designing rewards that capture more than final test correctness.
\newline \\
\textit{Challenges addressed: \ref{sec:c-long-horizon}, \ref{sec:c-ood-domains}}
\end{aicodeorangebox}

In the formal reasoning setting, we saw how ProofOptimizer leveraged RL by using the Lean environment to provide precise feedback for search and optimization. Designing environments like these for Code RL will bring the same type of verifiable signals to broader software engineering tasks.

\textbf{Collecting executable codebases}: Recent work in RLVR has shown success in solving algorithmic programming problems through DeepSeek-R1~\citep{deepseekai2025deepseekr1} and OpenAI o1. On SWE-Bench, SWE-RL \citep{wei2025swerladvancingllmreasoning} uses RL on a rule-based reward to improve performance. We find it promising to continue scaling the RL approach to problems collected from real-world software engineering repositories. These environments could improve reasoning, environment interaction, and tool usage by exposing agents to realistic edit-test-debug loops.

Several prior works~\citep{jain2024r2e, pan2024trainingsoftwareengineeringagents, guo2025syncmind, xie2025repost} curate executable environments for programming agents by supporting CI-based or heuristic repository installation. However, these efforts remain limited in scale and language coverage. Scaling them requires solving practical engineering problems: installing arbitrary repositories, storing executable images efficiently, starting containers quickly, and keeping environments reproducible as dependencies change. These infrastructure details are not incidental; without stable environments, training signals become noisy and benchmark results become difficult to interpret.

\textbf{Sourcing task prompts and rewards}: Beyond environments, large-scale reinforcement learning requires diverse tasks and reliable rewards. Prompts can be collected from GitHub~\citep{pan2024trainingsoftwareengineeringagents} or generated synthetically from repository states. Given executable repositories, we can source end-to-end problems beyond bug fixing, such as optimization, fuzzing, test generation, refactoring, and documentation repair. Tests provide a natural reward, but they are incomplete proxies: they can miss bugs, penalize acceptable alternative implementations, or encourage reward hacking~\citep{skalse2022defining}. Recent examples show that models can learn to bypass or exploit testing harnesses when optimized too directly against them~\citep{baker2025monitoring, denison2024sycophancy}.

\textbf{Rewards without execution}: Although execution is often the strongest signal, setting up execution environments can be expensive. Proxy metrics and learned judges may therefore be useful, especially when paired with symbolic signals. Earlier work used BLEU/CodeBLEU \citep{papineni2002bleu, ren2020codebleu} and BERTScore/CodeBERTScore \citep{zhang2019bertscore, zhou2023codebertscore} to assess text and code similarity. More code-aware approaches can incorporate syntax, data flow, or patch structure, as in Dolos \citep{maertens2022dolos} or edit-similarity rewards used in SWE-RL-style settings \citep{wei2025swerladvancingllmreasoning, ma2025sorft}. These signals should not replace execution when execution is available, but they can help shape behavior around code quality and semantic similarity.

\par\medskip\noindent\textbf{Adapting to Specialized and Quickly Changing Codebases}\phantomsection\label{sec:d-adaptation}\par\smallskip

\begin{aicodeorangebox}
Low-resource languages, custom APIs, library version updates, large codebases, and custom coding styles all surface the fact that code LMs struggle to adapt to unseen specialized contexts. Promising approaches include test-time adaptation, memory banks of codebase-specific information, and lightweight prompt or prefix tuning.
\newline \\
\textit{Challenges addressed: \ref{sec:c-global-understanding}, \ref{sec:c-ood-domains}}
\end{aicodeorangebox}

When humans are faced with an unfamiliar codebase, they spend time reading documentation, running examples, inspecting tests, and learning the local conventions. Models should have analogous adaptation mechanisms. Test-time training (TTT) adapts to a specific problem instance by training on a narrow set of in-distribution examples~\citep{akyurek2024surprisingeffectivenesstesttimetraining, sun19ttt}. In code, this could mean adapting to a repository, language, domain, or API version while retaining general coding knowledge. Synthetic examples can be generated in the specialized context and filtered with compilers or tests; agent trajectories can also be retained so that models learn from past failures rather than repeating them.

Retrieval and memory are a cheaper way to support this adaptation. A system can maintain a growing information bank of code, documentation, synthetic examples, previous attempts, and successful trajectories in the specialized context. Retrieving from this memory bank can prevent hallucinations of wrong APIs, improve style consistency, and expose the model to repository-specific invariants. Lightweight prompt tuning \citep{lester2021power} or prefix tuning \citep{li2021prefix} could also provide a way to represent codebase- or version-specific knowledge without full fine-tuning.

\subsubsection{Inference Time Approaches}

\par\medskip\noindent\textbf{Semantic-Aware Embeddings and Retrieval}\phantomsection\label{sec:directions-retrieval}\par\smallskip

\begin{aicodeorangebox}
In contrast to text, embeddings for code should incorporate execution and semantic information, improving retrieval. Beyond static retrieval, AI agents could dynamically navigate codebases using command-line tools, IDE functions, and program analysis.
\newline \\
\textit{Challenges addressed: \ref{sec:c-global-understanding}}
\end{aicodeorangebox}

\textbf{Semantic and execution-aware code embeddings}: When training LLMs, code is often treated as pure tokens rather than explicitly incorporating execution and semantics. As a result, code that is close in embedding space is often syntactically similar rather than semantically similar \citep{utpala2023language, zhao2023get}. Before the LLM era, however, there were efforts to learn program representations that encode semantics, including neural modules for program operations \citep{nye2020representing} and execution-aware latent representations for partial or full programs \citep{zohar2018automatic, ellis2019write, chen2021latent}. Revisiting these ideas may improve retrieval and make code representations more aligned with program behavior.

\textbf{Better retrieval-augmented code generation}: Retrieval is especially useful in specialized contexts where the model has not seen enough examples to fully understand a language, API, or codebase. Documentation, function definitions, and local usage examples can teach syntax, available abstractions, and conventions. However, retrieving relevant code is only half the problem. The model must also reuse the retrieval correctly, adapting algorithms, translating between languages, and recognizing subtle differences between the retrieved context and the current task. This suggests that retrieval-augmented code models should be trained not only to consume retrieved snippets, but also to reason about how those snippets should be modified.

\textbf{Retrieving via code navigation on the fly}: Static retrieval indexes can be expensive to build and can become stale as a codebase evolves. Instead of relying only on embeddings, agents can retrieve information by navigating the codebase directly, using commands such as \texttt{cd}, \texttt{ls}, and \texttt{grep}, IDE actions such as jumping to definitions or finding references, and static-analysis tools that expose ASTs or dependency graphs. This frames retrieval as an interactive code-understanding task rather than a one-shot nearest-neighbor lookup.

\par\medskip\noindent\textbf{Incorporating SWE Tools}\phantomsection\label{d:sec-agents-tools}\par\smallskip

\begin{aicodeorangebox}
Software engineers integrate a variety of domain-specific tools when writing code. By repeatedly interacting with tools in an RL-style manner, AI systems can learn when to invoke tools, how to invoke them, and how to incorporate their outputs into the next step of reasoning.
\newline \\
\textit{Challenges addressed: \ref{sec:c-tool-use}}
\end{aicodeorangebox}

As mentioned in Sec.~\ref{sec:c-tool-use}, SWE agents should understand the intricacies of programming tools and be able to invoke them autonomously when needed. There are three skills to learn: which tool to use, how to use the tool, and how to incorporate the result. Similar to how models learn to play complicated games, intelligent tool integration may be learned through repeated interactions with tools in an RL-style manner. First, the interface of the tool must be precisely specified. Next, data containing repeated interactions with the tool should be collected. Finally, multiple rounds of RL and expert iteration can improve the model's understanding of successful and unsuccessful tool use.

\textbf{Neurosymbolic approaches}: Code is a unique domain because there is a vast body of programming-languages research to build on, including abstract interpretation \citep{cousot1977abstract}, concolic testing \citep{godefroid2005dart, sen2005cute}, model checking \citep{clarke1997model}, linting, and type-checking \citep{cardelli1996type}. These tools can surface bugs, enforce invariants, narrow search spaces, and provide structured feedback. Traditional PL tools often require precise specifications, can be computationally expensive, and may produce false positives. LLMs can help mitigate these issues by proposing specifications, localizing likely errors, and deciding when a symbolic tool is useful.

LINC and ProofOptimizer highlight this broader pattern. In LINC, the model
translated natural language reasoning into executable symbolic programs. In ProofOptimizer,
the model proposed proof transformations while the compiler ensured correctness.
Neurosymbolic methods potentially allow AI systems to explore a large search space while maintaining guarantees throughout.

We provide a few examples of this potential integration. During code generation, program-analysis techniques could be applied to shorter snippets of AI-generated code to surface bugs or prove properties. To improve code understanding, LLMs can be trained with program structure such as abstract syntax trees \citep{gong2024ast}. When debugging a large codebase, AI could first narrow down the suspicious region before handing the problem to more precise symbolic tools. During code generation in DSLs, LLMs can leverage grammars for constrained decoding~\citep{poesia2022synchromesh, geng2023grammar, wei2023copiloting}. During refactoring, static analysis can identify whether new errors have been introduced and cut off unpromising search paths.

Taken together, these directions suggest a common principle: the next stage of AI for code should be organized around feedback-rich interaction with programs. Better data gives models traces of how software behaves; RL environments turn those traces into learning signals; adaptation and retrieval help models understand local codebases; and tool use grounds model reasoning in executable evidence.
