\subsection{Discussion}
In this chapter, we first briefly describe several tasks 
that we believe should be considered when it comes to AI for software engineering,
focusing on formal verification as one such task. 
We have also highlighted a few cross-cutting challenges that permeate throughout many software-engineering tasks. 
Finally, to drive progress in AI for code, we've pinpointed several promising research directions for 
alleviating these challenges and advancing AI towards being a more capable software engineer.

In many ways, the lessons developed throughout this thesis were used in developing this roadmap, but more broadly
can guide future benchmark research and develop analogous roadmaps in other fields. 
In Chapter 2, we saw how each metric revealed a different slice of AI systems and showed different insights. 
When it comes to benchmark research, the first step is deciding what to measure. In this roadmap, we
decided on our set of tasks after looking at the skills that software engineers needed to have and compared it with
the skills that today's coding agents have. We also thought more broadly about the longer-term horizon. For example,
as AI generates more code, it will be more important for there to be correctness guarantees, which motivated tasks such as
software testing, program analysis, and formal verification.

In Chapter 3, we showed the importance of deeper analysis, which we used to form the set of challenges discussed in the roadmap. 
We analyzed both benchmarks and real-world use cases guided by the tasks in the previous section. For example, by analyzing how a model fails tasks in SWE-Bench, we were able to
find that models sometimes struggled at interpreting vague specifications the way that human programmers would. By trying today's
frontier coding agents on a set of out-of-distribution tasks, we were able to see how they struggled and failed. Overall, a section like this 
is created through looking at the tasks in the previous section, collecting intuition about failure modes, and cleanly categorizing them
into a set of modular issues and challenges.

Finally, Chapter 4 showed how evaluations can shape the direction of future research and science. In our roadmap paper, we designed this section
by first examining both the tasks and the challenges in the previous sections. Then, we looked at the literature for techniques addressing each of the
challenges, and synthesized both these and our own ideas into a collection of paths forward. Many of these techniques are inspired by benchmarks:
for example, CRUXEval inspired a family of work training using execution feedback. These works inspired a section in our
paths forward section about augmenting code data with program information: starting with execution, but also including 
program instrumentation, dynamic analysis, and beyond.

Overall, in order to design a roadmap, I believe we must first think critically about where a field is headed: not just in the near-term,
but rather a longer-term vision about what objectives and metrics will be important a few years out. Then, we need to think about the core 
challenges today that are preventing us from achieving this vision. Finally, we should reflect on the research trajectory that the field has been
on, build intuition for what seems promising, and use that to design a set of paths forward. Of course, there is no way to precisely predict
the future: only time and experimentation will tell which directions prevail at the end. However, roadmaps can guide a community towards a specific 
set of visions and goals, helping everyone better understand the state of a field. 
